loading/cleaning/wrangling/exploring

### 2024-07-05 cdcdata-exercise/deathcount_cleaned.qmd
updated R code 07.05.24

```{r}
library(readxl)
#update.packages("tidyverse")
library(tidyverse)
#update.packages("dplyr")
library(dplyr)
#update.packages('mgcv')
library(mgcv)
```

@@

```{r}
head(DthCount_data)
```

```{r}
str(DthCount_data)
```

@@

```{r}
# Assuming DthCount_cleaned is your cleaned data frame
colnames(DthCount_cleaned) <- c("Jurisdiction", "Week_Ending_Date", "State", "Year", "Week","Cause","Death_Count","Cause_Subgroup","Period","Suppress","Note","Avg_No_Deaths","Variance_2015_2019_2020", "Perc_Diff_2015_2019_2020","Types")
```

@@

```{r}
# Write the cleaned data to a new CSV file
write.csv(DthCount2_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DthCount2_cleaned.csv", row.names = FALSE)
```

```{r}
DthCount3_cleaned <- DthCount2_cleaned %>% select(-Suppress,-Note,-Types)

@@

```{r}
# Write the cleaned data to a new CSV file
write.csv(DthCount3_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DthCount3_cleaned.csv", row.names = FALSE)
```

```{r}
# Assuming dat_cleaned is your data frame
DthCount3A_cleaned <- DthCount3_cleaned[apply(DthCount3_cleaned, 1, function(x) !all(is.na(x))), ]

# Print the modified data frame to verify the changes
head(DthCount3_cleaned)

@@

```{r}
# Convert the 'state' column to a factor
DthCount3A_cleaned$State <- as.factor(DthCount3_cleaned$State)

# Display the resulting data frame
head(DthCount3_cleaned)
```

@@

@@

```{r}

library(tidyr)

DeathCnt_data <- DthCount3A_cleaned %>% drop_na()

# Verify that NAs have been removed
glimpse(DeathCnt_data)
```

```{r, message=FALSE, warning=FALSE}
## Example of a bivariate plot between 'state' and 'varroa_mites'
ggplot(DeathCnt_data, aes(x = State, y = Death_Count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Death Count by State",
       x = "State",
       y = "Death Count")
```

@@

```{r}
# Write the cleaned data to a new CSV file
write.csv(DeathCntFinal_data, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DeathCntFinal_data.csv", row.names = FALSE)
```

################################Eddie Cruz #################################################################

@@

```{r}
# Handle missing values (for simplicity, we'll remove rows with any NAs)
data <- na.omit(data)

@@

```{r}
# Remove the original Week_Ending_Date column to avoid issues with high levels
data$Week_Ending_Date <- NULL

@@

# Save the synthetic dataset to a CSV file
write.csv(synthetic_data, "synthetic_data.csv", row.names = FALSE)
```

>>>>

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(mgcv)
```

@@

```{r}
#Review variables
head(DthCount_data)
```

```{r}
# Type of variable
str(DthCount_data)

@@

```{r}
# Clean up variable labels
colnames(DthCount_cleaned) <- c("Jurisdiction", "Week_Ending_Date", "State", "Year", "Week","Cause","Death_Count","Cause_Subgroup","Period","Suppress","Note","Avg_No_Deaths","Variance_2015_2019_2020", "Perc_Diff_2015_2019_2020","Types")
```

@@

```{r}
# Write the cleaned data to a new CSV file
#write.csv(DthCount2_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II #Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DthCount2_cleaned.csv", row.names = FALSE)
```

```{r}
#suppress three variables
DthCount3_cleaned <- DthCount2_cleaned %>% select(-Suppress,-Note,-Types)
```

@@

```{r}
# Write the cleaned data to a new CSV file
#write.csv(DthCount3_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II #Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DthCount3_cleaned.csv", row.names = FALSE)
```

```{r}
# Assuming dat_cleaned is your data frame
DthCount3A_cleaned <- DthCount3_cleaned[apply(DthCount3_cleaned, 1, function(x) !all(is.na(x))), ]
```

```{r}
# Print the modified data frame to verify the changes
head(DthCount3_cleaned)

@@

```{r}
# Convert the 'state' column to a factor
DthCount3A_cleaned$State <- as.factor(DthCount3_cleaned$State)
```

```{r}
# Display the resulting data frame
head(DthCount3_cleaned)
```

@@

```{r}
str(DthCount3A_cleaned)
```
##########################################################Edward Cruz###########################################################################
```{r}
library(ggplot2)

@@

```{r}
#Remove NA values
DeathCnt_data <- DthCount3A_cleaned %>% drop_na()
```

```{r}
# Verify that NAs have been removed
glimpse(DeathCnt_data)
```
# View a cleaner plot
```{r, message=FALSE, warning=FALSE}
## Example of a bivariate plot between 'state' and 'varroa_mites'
ggplot(DeathCnt_data, aes(x = State, y = Death_Count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Death Count by State",
       x = "State",
       y = "Death Count")
```

@@

```{r}
# Write the cleaned data to a new CSV file which will be used to create synthetic data
write.csv(DeathCntFinal_data, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DeathCntFinal_data.csv", row.names = FALSE)
```

################################Eddie Cruz & Leonels Salazar #################################################################

@@

```{r}
# Handle missing values (for simplicity, we'll remove rows with any NAs)
data <- na.omit(data)
```

@@

```{r}
# Remove the original Week_Ending_Date column to avoid issues with high levels
data$Week_Ending_Date <- NULL
```

@@

# Save the synthetic dataset to a CSV file
write.csv(synthetic_data, "synthetic_data.csv", row.names = FALSE)
```
#####################################################Leonel Salazar#############################################################################
